Tidal streams in triaxial systems

### 2015-02-19 ms.tex
hyperref
intro end

% packages
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{bbold}
\usepackage{arydshln}
\usepackage{algpseudocode}
\renewcommand\algorithmicthen{}
\renewcommand\algorithmicdo{}
\usepackage{mathrsfs}
\usepackage{subfigure}

@@

@@

For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced. As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed. Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}. However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system. Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction in Section~\ref{sec:}). This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{}) or frequency diffusion time (Section~\ref{}). In Section~\ref{} we look at how these relate to the the mixing rate of small ensembles of orbits.

[summarizing paragraph connecting everything, mention mixing, relation to tidal streams]

\section{Numerical methods}

@@

\subsection{Lyapunov exponents}

@@

\subsection{Numerical estimation of fundamental frequencies}

@@

\subsection{Test particle ensembles}

@@

\section{Experiments}

@@

[Lyapunov exponent grids...]

@@

\acknowledgements
The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, \emph{Stream Team}, etc...
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.

>>>>

% packages
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{bbold}
\usepackage{arydshln}
\usepackage{algpseudocode}
\renewcommand\algorithmicthen{}
\renewcommand\algorithmicdo{}
\usepackage{mathrsfs}
\usepackage{subfigure}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}

@@

%TODO:
%The classical way to detect chaotic motion is
%the computation of the Lyapunov exponents. We
%claim that for weakly chaotic solutions, the frequency
%analysis gives information more rapidly
%and more precisely. The reason is that the frequency
%analysis is an instantaneous observation
%of the fundamental frequencies which are essential
%parameters of the dynamical system, while
%the usual computation of Lyapunov exponents
%acts like a mean and needs a very long time span
%to provide reliable results (for references and
%discussion on Lyapunov exponents, see refs. [3,
%12].

@@

For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced. As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed. Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}. However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system. Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction with the experiments of Section~\ref{sec:experiments}). This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).

These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit. For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}). The evolution of tidal debris will depend on the ensemble mixing rate of the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}. Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is low, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.

This paper is organized as follows: In Section~\ref{sec:methods}, we outline the numerical methods used for the experiments of Section~\ref{sec:experiments}. In Section~\ref{sec:discussion}, we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.

\section{Numerical methods}\label{sec:methods}

@@

\subsection{Lyapunov exponents} \label{sec:lyap}

@@

\subsection{Numerical estimation of fundamental frequencies}\label{sec:naff}

@@

\subsection{Test particle ensembles} \label{sec:ensemble}

@@

\section{Experiments}\label{sec:experiments}

@@

\begin{figure}[!h]
\centering
	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_yz.png}
	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
	\label{fig:lyapmap}
\end{figure}

We also compute the Lyapunov exponents for all orbits in the grid by integrating their orbits for 250 orbital periods. Figure~\ref{fig:lyapmap} shows the initial condition grids again, but now the greyscale intensity corresponds to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$. These Lyapunov exponent maps show that, for a fixed integration time, only the most chaotic orbits are identified. NAFF is able to detect small differences in the frequencies and thus measure small diffusion rates with comparable integration times.

@@

\acknowledgements
The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.
